Graph bipartitioning and you

### 2016-01-30 evolutionaryComputation/2-practical/report.tex
added student ttest and cleared some other stuff up
adding more text and both tables

While implementing ILS, I had it turn out results much worse than MLS.
The initial implmentation would just stop if it found a worse local optimum.
The description sais it should go back to the previous optimum. So I guess
that you should try again after that? But then I wonder how often you should
try again.

@@

I don't think there is a way to do it any better without drastically making
the algorithm more complex.
This current search for improvement was just based on the fact that ILS was
underperforming to MLS, which shouldn't be the case according to my memory
from the classes. Now its overperforming quite drastically so I'm happy.

I implemented GLS with the evolution class. The only thing I had to do really
was write the crossover method, the rest was already functional.

@@

@@

@@

\section{Results}
\subsection{Time}
\pgfplotstableread{runtimes.table}\dataTable
\begin{tikzpicture}
\pgfplotsset{set layers}
\begin{axis}[
	ybar stacked,
	xlabel={Algorithm},
	xticklabels from table={runtimes.table}{name},
	xtick=data,
	x tick label style={rotate=90,anchor=east},
	ylabel={Time}
]

% Layers needed to get error bars on top of lower blue rectangle
\addplot[color=white]
table[on layer=axis grid, x expr=\coordindex, y=firsty]{\dataTable};

@@

\input{runtimes.textable}

\subsection{Optima}
\pgfplotstableread{results.table}\dataTable
\begin{tikzpicture}
\pgfplotsset{set layers}
\begin{axis}[
	ybar stacked,
	xlabel={Algorithm},
	xticklabels from table={results.table}{name},
	xtick=data,
	x tick label style={rotate=90,anchor=east},
	ylabel={Fitness (less is better)}
]
% Layers needed to get error bars on top of lower blue rectangle
\addplot[color=white]
table[on layer=axis grid, x expr=\coordindex, y=firsty]{\dataTable};

@@

\end{tikzpicture}
\\
\input{results.textable}
\end{empfile}
\end{document}

>>>>

\subsection{MLS}
MLS basically involved implementing the initial vertex swap first improvement
local search and tying it into the existing framework. I did this by just
using the existing createPopMethods with a new solution generator. Then
I just mapped this population with the local search (so the type is
$IMember \to IMember$, but the solution changes).

\subsection{ILS}
While implementing ILS, I had it turn out results much worse than MLS.
The initial implmentation would just stop if it found a worse local optimum.
The description sais it should go back to the previous optimum. So I guess
that you should try again after that? But then I wonder how often you should
try again.

@@

I don't think there is a way to do it any better without drastically making
the algorithm more complex.
This current search for improvement was just based on the fact that ILS was
underperforming to MLS, which shouldn't be the case according to my memory
from class. Now its overperforming quite drastically so I'm happy.

\subsection{GLS}
I implemented GLS with the evolution class. The only thing I had to do really
was write the crossover method, the rest was already functional.

@@

Here I also started to think about my own topic to investigate, at first
I just wanted to do varying population sizes and varying selection methods.
But I thinks this takes a little to much time and it doesn't quite fit
on the tTest table so I decided the varying popsizes (spoilers: it produces
better solution at the cost of exponentially more time).

@@

\subsubsection{Graph pruning}
I decided to use a prune graph to speed up the algortihm. In the beginning I
create 2 graphs. One normal and one pruned, The pruning works by selecting
comparing all connections against the id, if lower we ignore the connection.
To make sure we only do this on doubly linked graphs 2 graphs are created.
If the prune graph has exactly half the score of a random solution we assume
its a doubly linked graph so we can prune it.
The pruning effectivally halfs the ammount of connections that need to be
checked by each itteration (this happens in the evaluation). The
FidduciaMathesis has even more gain because it has a lot of book keeping to
do with the connections (although it tries to reduce this with the freecell
array, having a smaller initial graph is even better).

\subsection{Fidducia Mathesis}
I was somewhat sceptical of my local search implementation. I had quite 
different results from my classmates so to compare I decided to implement
the fidducia mathesis local search to. It was quite a lot of work but it
truly is significantly faster.

Finding the paper was somewhat difficult as the original was behind a paywal.
But somoene put a copy on github so I just used that one. The algorithm 
involves a lot of cleaver book keeping:

@@

\section{Results}
\subsection{Fiduccia Mathesis}
\subsubsection{Time}
\pgfplotstableread{fidRuntimes.table}\dataTable
\begin{tikzpicture}
\pgfplotsset{set layers}
\begin{axis}[
	ybar stacked,
	xlabel={Algorithm},
	xticklabels from table={fidRuntimes.table}{name},
	xtick=data,
	x tick label style={rotate=90,anchor=east},
	ylabel={Time}
]
% Layers needed to get error bars on top of lower blue rectangle
\addplot[color=white]
table[on layer=axis grid, x expr=\coordindex, y=firsty]{\dataTable};

@@

\input{fidRuntimes.textable}

\subsubsection{Optima}
\pgfplotstableread{fidResults.table}\dataTable
\begin{tikzpicture}
\pgfplotsset{set layers}
\begin{axis}[
	ybar stacked,
	xlabel={Algorithm},
	xticklabels from table={fidResults.table}{name},
	xtick=data,
	x tick label style={rotate=90,anchor=east},
	ylabel={Fitness (less is better)}
]
% Layers needed to get error bars on top of lower blue rectangle
\addplot[color=white]
table[on layer=axis grid, x expr=\coordindex, y=firsty]{\dataTable};

@@

\end{tikzpicture}
\\
\input{fidResults.textable}

\subsection{Vertex swap first improvemnt}
\subsubsection{Time}
\pgfplotstableread{vertRuntimes.table}\dataTable
\begin{tikzpicture}
\pgfplotsset{set layers}
\begin{axis}[
	ybar stacked,
	xlabel={Algorithm},
	xticklabels from table={vertRuntimes.table}{name},
	xtick=data,
	x tick label style={rotate=90,anchor=east},
	ylabel={Time}
]
% Layers needed to get error bars on top of lower blue rectangle
\addplot[color=white]
table[on layer=axis grid, x expr=\coordindex, y=firsty]{\dataTable};

\input{vertRuntimes.textable}

\subsubsection{Optima}
\pgfplotstableread{vertResults.table}\dataTable
\begin{tikzpicture}
\pgfplotsset{set layers}
\begin{axis}[
	ybar stacked,
	xlabel={Algorithm},
	xticklabels from table={vertResults.table}{name},
	xtick=data,
	x tick label style={rotate=90,anchor=east},
	ylabel={Fitness (less is better)}
]
% Layers needed to get error bars on top of lower blue rectangle
\addplot[color=white]
table[on layer=axis grid, x expr=\coordindex, y=firsty]{\dataTable};

\end{tikzpicture}
\\
\input{vertResults.textable}
\end{empfile}
\end{document}
